### 2009-04-01 docs/eos_notes/eos_notes.tex
new section on eos calls and temperature usage

\section{EOS Calls}
\section{Temperature Usage}