### 2010-04-02 docs/preface/preface.tex
font play

MAESTRO is developed at the Center for Computational Sciences and
Engineering at Lawrence Berkeley National Laboratory and at Stony
Brook University.  Development was supported by the SciDAC Program of
the DOE Office of Mathematics, Information, and Computational Sciences
under the U.S. Department of Energy under contract
No.\ DE-AC02-05CH11231 to LBL and by a DOE/Office of Nuclear Physics
Outstanding Junior Investigator award, grant No.\ DE-FG02-06ER41448,
to Stony Brook. \\

\noindent The current version of the MAESTRO User's Guide can be found in 
the MAESTRO source tree under {\tt fParallel/MAESTRO/docs/}.