### 2010-10-05 docs/beta0/beta0.tex
start of a document to describe our understanding of beta_0

The goal of $\beta_0$ is to capture the expansion of a displaced fluid
element due to the stratification of the atmosphere.  MAESTRO computes
$\beta_0$ as:
\begin{equation}
\beta_0(r,t) = \rho_0 \exp\left (  \int_0^r  \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial r^\prime} dr^\prime \right )
\end{equation}

\section{Constant Composition}
Consider an isentropically stratified atmosphere, with a constant
composition as a function of $r$.  If you displace a parcel of fluid,
it will expand adabatically and continue to rise until its density
matches the background density.  
Even if $\gammabar$ is not constant in $r$, following from the
definition of $\beta_0$,
\begin{equation}
\frac{1}{\beta_0} \frac{d\beta_0}{dr} = \frac{1}{\gammabar p_0} \frac{dp_0}{dr}
\end{equation}
and the definition of $\Gamma_1$,
\begin{equation}
\Gamma_1 = \left . \frac{d \log p}{d \log \rho} \right |_s
\end{equation}
So, at constant entropy, from the definition of $\Gamma_1$, it must hold
that
\begin{equation}
\frac{1}{\rho} \frac{d \rho}{dr} = \frac{1}{\Gamma_1 p} \frac{d p}{dz} \enskip .
\end{equation}
Comparing to the defintion of $\beta_0$ then
\begin{equation}
\frac{1}{\beta_0} \frac{\beta_0}{dr} =\frac{1}{\gammabar p_0}\frac{dp_0}{dr} = \frac{1}{\rho_0} \frac{d\rho_0}{dr}  \enskip .
\end{equation}
Therefore, $\beta_0 = \rho_0$.

This means that if we have a constant composition and an
isentropically stratified atmosphere, as we displace a fluid element,
it will always remain neutrally buoyant.

\section{Composition Gradient}