### 2011-04-03 docs/enthalpy/enthalpy.tex
Added more stuff about derivation of constraint.

\section{Evolution Equations}

The compressible and low Mach number formulations of the governing
equations both share the unapproximated continuity and momentum equations
shown here (where $p = p_0 + \pi$).
\begin{eqnarray}
\frac{\partial(\rho X_k)}{\partial t} &=& -\nabla\cdot(\rho X_k\Ub) + 
\rho\omegadot_k,\label{eq:species}\\
\frac{\partial\Ub}{\partial t} &=& -\Ub\cdot\nabla\Ub  - 
  \frac{1}{\rho}\nabla\pi - 
  \frac{\rho-\rho_0}{\rho} g\eb_r,\label{eq:momentum}
\end{eqnarray}

In the compressible formulation we complete the system with an energy
equation as well as an equation of state; in the low Mach number
formulation we can derive a constraint on the velocity by 
setting $p_{EOS} = p(\rho,h,X_k) = p_0$ and
differentiating the equation of state along particle paths.
In this case adding the energy equation would over-constrain the system, 
but its solution can help us in providing a diagnostic capability for the
solution.  We can write the energy equation in terms of internal energy, $e$,
or enthalpy, $h$; these two equations are analytically equivalent.
\begin{eqnarray}
\frac{\partial(\rho h)}{\partial t} + \nabla\cdot(\rho h\Ub) 
 &=& \frac{Dp}{Dt} + \rho\Hnuc \nonumber \\
%
\frac{\partial(\rho e)}{\partial t} + \nabla\cdot(\rho e \Ub)
 &=& - p\nabla\cdot \Ub + \rho\Hnuc \nonumber 
\end{eqnarray}
In low Mach number combustion, $Dp/Dt = 0,$ which makes the enthalpy equation
preferable to the internal energy equation because we don't need to
evaluate $p \nabla \cdot \Ub.$

\section{Derivation of Constraint Using Enthalpy Equation}

The velocity constraint equation was constructed by differentiating 
the pressure along particle paths, using the equation of state.  
Specifically, we wrote $p_{EOS} = p(\rho,T,X_k) = p_0$
and expressed $DT/Dt$ in terms of $Dh/Dt:$
\begin{equation}
\frac{DT}{Dt} = \frac{1}{\rho c_p} \left( (1 - \rho h_p) \frac{D p}{D t}
- \sum_k \rho \xi_k \omegadot_k + \rho H \right) \enskip , \label{eq:dTdt}
\end{equation}
where $c_p = \left.\partial h/\partial T\right|_{p,X_k}$ is the
specific heat at constant pressure,
$\xi_k = \left.\partial h/\partial X_k \right|_{p,T}$,
and $h_p = \left.\partial h/\partial p\right|_{T,X_k}$
for convenience.

Then we derived
\begin{eqnarray*}
\nabla \cdot \ubold &=& \frac{1}{\rho p_\rho} \left(
- \frac{D p}{D t} + \frac{p_T}{\rho c_p}
  \left( (1 - \rho h_p) \frac{D p}{D t} - \rho \sum_k \xi_k \omegadot_k + \rho H \right)
+ \sum_k p_{X_k} \omegadot_k \right)  \\
                 &=& \frac{1}{\rho p_\rho}
  \left( \frac{p_T}{\rho c_p}(1  - \rho h_p) - 1 \right) \frac{D p}{D t}
 + \frac{1}{\rho p_\rho} \left(
  \frac{p_T}{\rho c_p} (\rho H - \rho \sum_k \xi_k   \omegadot_k)
                               + \sum_k p_{X_k} \omegadot_k \right) \enskip .
\end{eqnarray*}
When we derived this expression we explicitly retained the dependence of $h$ on $p$,
as shown by the presence of the $h_p$ term.

Then, replacing $p$ by $p_0(r)$, $Dp/Dt$ becomes $\ubold \cdot
\nabla p_0$, and, recalling the definition of $H$, the divergence constraint can
be written
\begin{equation}
\nabla \cdot \ubold + \alpha \ubold \cdot \nabla p_0 =
\frac{1}{\rho p_\rho} \left(
   \frac{p_T}{\rho c_p} \left(\nabla \cdot (\kappa \nabla T)
  - \sum_k\rho (q_k + \xi_k) \omegadot_k \right)
 + \sum_k p_{X_k} \omegadot_k \right)  \equiv \tilde{S} \enskip \label{eq:full_divu_constraint} ,
\end{equation}
where we define
\begin{equation}
\alpha(\rho,T) \equiv - \left( \frac{(1 - \rho h_p )p_T - \rho c_p}{\rho^2
  c_p p_\rho} \right) \enskip . \label{eq:alphadef}
\end{equation}

\section{Derivation of Constraint Using Energy Equation}

We could also derive the constraint by again differentiating $p_{EOS} = p(\rho,T,X_k) = p_0$
but this time writing $DT/Dt$ in terms of $De/Dt:$

\begin{equation}
\frac{DT}{Dt} = \frac{1}{\rho c_v} \left( -e_p \frac{D p}{D t}
- \sum_k e_k \frac{D X_k}{D t} - p \nabla \cdot U 
+ \rho H \right) \enskip , \label{eq:dTdt}
\end{equation}
where $c_v = \left.\partial e/\partial T\right|_{p,X_k}$ is the
specific heat at constant volume,
and $e_p = \left.\partial e/\partial p\right|_{T,X_k}$
for convenience.

Then...

\section{Enthalpy vs Energy Equation}

The full enthalpy equation, with no approximations, appears as:
\begin{equation}
\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) + 
  \frac{Dp}{Dt} + \rho\Hnuc \label{eq:enthalpy}
\end{equation}
Here, $h = e + p/\rho$ is the specific enthalpy, with $e$ the specific
internal energy.  In the low Mach number formulation, we replace $p$
with $p_0$ in the $Dp/Dt$ term, however, the definition of enthalpy
implicitly contains a pressure.  When calling the equation of state,
we take $h$ and $\rho$ as inputs.  The equation of state is expressed
in terms of $T$ and $\rho$, so it iterates until it finds the $h$ that
we desire.  This $h$ will be of the form $h = e + p_\mathrm{EOS}/\rho$,
where $p_\mathrm{EOS}$ is the pressure returned from the EOS.  Note that
$p_\mathrm{EOS}$ may not be equal to $p_0$---this may be what
causes us to drive off of the constraint.

The mismatch between the pressure implicit in the definition of $h$
and $p_0$ can be seen by substituting $h = e + p/\rho$ into the
enthalpy equation, where we replace $p$ with $p_0$ in the $Dp/Dt$ term:
\begin{eqnarray}
\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) + 
  \frac{Dp_0}{Dt} + \rho\Hnuc \nonumber \\
%
\frac{\partial(\rho e)}{\partial t} + \frac{\partial p}{\partial t} &=&
 -\nabla\cdot(\rho e\Ub) -\nabla\cdot(p\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc \nonumber \\
%
\frac{\partial(\rho e)}{\partial t} &=&
 -\nabla\cdot(\rho e\Ub) - p\nabla\cdot\Ub + \rho\Hnuc + 
  \left \{ \frac{Dp_0}{Dt} - \frac{Dp}{Dt} \right \} \nonumber 
\end{eqnarray}
We note that the last equation appears to be an internal energy
evolution equation, but with an additional forcing term: $D(p_0 - p)/Dt$.

\section{Outstanding Questions}

\begin{enumerate}
\item Why do we want to start with enthalpy instead of internal energy?

   We believe that the original desire stems from our experience with smallscale 
   combustion.  There, stratification is not important and $Dp_0/Dt = 0$, so
   the enthalpy equation becomes a conservation equation for $(\rho h)$.

\item We substitute $p$ for $p_0$ in two places: 1) in the derivation of the
constraint, 2) in the enthalpy evolution equation.   Are these substitutions
consistent with each other?  If they are not consistent, which of these 
substitutions is responsible for filtering the acoustics from the system?

\item In the EOS, we take $h$ and construct $e$ by subtracting off
  $p$---depending on the inputs to the EOS, we treat the `$p$' part of $h$
  differently.

   if we come into the EOS with $\rho$, $T$, then we get $p$ and $e$ from
   the EOS and compute $h$.

   if we come into the EOS with $\rho$, $h$, then we find a $T$ such that
   $e + p(\rho,T)/\rho = h$, where $p(\rho,T)$ is the pressure from the EOS.
   It is not necessarily equal to $p_0$.

   if we come into the EOS with $\rho$, $p_0$, then find a $T$ that matches
   $p_0$, and then $h = e + p_0/\rho$.  In this case, there is no mismatch
   in the enthalpy equation.

\item do we want to always insist that the `$p$' hidden in the definition of
   $h$ is $p_0$?

\end{enumerate}