### 2011-06-14 docs/compilers/compilers.tex
start some documentation on compilers

\section{Compilers}

Here, we list notes on problems with specific compilers and Maestro.

\subsection{Intel}

Overall, these tend to be flaky.

\subsection{gfortran}

Probably the best compilers for Maestro.  Versions 4.5 and 4.6 seem to work
really well and the debugging features are very reliable.

\subsection{PGI}

(on Kraken) The bounds checking routines do not like the way some of
the arrays are allocated and flag them as errors.  For instance, in
{\tt layout.f90}, we allocate a box as:
\begin{verbatim}
    type(box) :: bxs(3**get_dim(b))
\end{verbatim}
but the bounds checking routine seems to fail on this.

(on hopper) There are problems with code compiled with {\tt -fast}
instead of just {\tt -O}.  gcc/gfortran seem to work better and 
are only slighly slower.

\subsection{PathScale}

These are the prefered compilers on jaguar.