### 2011-06-23 docs/pert_notes/paper.tex
use maestrosymbols.tex

\documentclass[11pt]{article}

\tolerance=600

\usepackage{amsmath,color,pdflscape}

% Margins
\usepackage[lmargin=0.5in,rmargin=1.5in,tmargin=1.0in,bmargin=1.0in]{geometry}

\input ../maestrosymbols

\title{Notes on Prediction Types}

\setlength{\marginparwidth}{1.0in}
\newcommand{\MarginPar}[1]{\marginpar{%
\vskip-\baselineskip %raise the marginpar a bit
\raggedright\tiny\sffamily
\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}

\begin{document}

\maketitle
\tableofcontents
\cleardoublepage

\input pert

\end{document}