### 2011-12-13 docs/compilers/compilers.tex
update the compiler folklore

\section{Compilers}

Here, we list notes on problems with specific compilers and \maestro.

\subsection{Intel}

Overall, these tend to be flaky.

\subsection{gfortran}

Probably the best compilers for \maestro.  Versions 4.5 and 4.6 seem to work
really well and the debugging features are very reliable.

\subsection{PGI}

The general feeling with these compilers is that they produce slower
code than the PathScale compilers.

\ \\
Machine-specific notes:
\begin{itemize}

\item (on Kraken) The bounds checking routines do not like the way some of
the arrays are allocated and flag them as errors.  For instance, in
{\tt layout.f90}, we allocate a box as:
\begin{verbatim}
    type(box) :: bxs(3**get_dim(b))
\end{verbatim}
but the bounds checking routine seems to fail on this.

\item (on hopper) There are problems with code compiled with {\tt -fast}
instead of just {\tt -O}.  gcc/gfortran seem to work better and 
are only slighly slower.
\end{itemize}

\subsection{PathScale}

These are the prefered compilers on jaguar, however, the PathScale
compilers are no longer actively supported on the Crays and therefore
should be considered depricated.

\ \\
Machine specific notes:
\begin{itemize}

\item  To run with BDS, change {\tt -Ofast} to {\tt -O2} in {\tt Linux\_pathscale.mak}

\item (on jaguar) with OpenMP, the Pathscale compiler warns about
  variables that are shared, but doesn't always give useful line
  numbers.  In debug mode, add ``-fullwarn'' to the compilation to get
  the compiler output to tell you where the problems are.

\end{itemize}

\subsection{Cray}

So far, no luck with these compilers (versions 7.4.3 and 8.0.0.135).