### 2011-12-22 docs/compilers/compilers.tex
add information regarding the IBM compilers

\section{Compilers}

Here, we list notes on problems with specific compilers and \maestro.

%-----------------------------------------------------------------------------
\subsection{Cray}

So far, no luck with these compilers (versions 7.4.3 and 8.0.0.135).

%-----------------------------------------------------------------------------
\subsection{gfortran}

Probably the best compilers for \maestro.  Versions 4.5 and 4.6 seem to work
really well and the debugging features are very reliable.

%-----------------------------------------------------------------------------
\subsection{IBM (xl* compilers)}

Recent versions of these compilers seem to work ok.  Some notes:
\begin{itemize}
\item You need to use {\tt xlf95}, not {\tt xlf90} to compile, since
  MAESTRO relies on the Fortran 95 feature that arrays are
  automatically deallocated when you leave a subroutine.

\item Many (all?) IBM machines have a different endian-ness than
  Intel-like chips.  \boxlib\ plotfiles store information used to
  dechipher the endian-ness of the data.  To do this properly on
  the IBM machines, you need to set the preprocessor flag {\tt BL\_AIX},
  which is interpretted by {\tt fabio\_c.c}.  This can be done for the
  IBM compilers by adding {\tt -Wp,-DBL\_AIX} to the {\tt CFLAGS} variable.

\end{itemize}

%-----------------------------------------------------------------------------
\subsection{Intel}

Overall, these tend to be flaky.

%-----------------------------------------------------------------------------
\subsection{PathScale}

These are the prefered compilers on jaguar, however, the PathScale
compilers are no longer actively supported on the Crays and therefore
should be considered depricated.

\ \\
Machine specific notes:
\begin{itemize}

\item  To run with BDS, change {\tt -Ofast} to {\tt -O2} in {\tt Linux\_pathscale.mak}

\item (on jaguar) with OpenMP, the Pathscale compiler warns about
  variables that are shared, but doesn't always give useful line
  numbers.  In debug mode, add ``-fullwarn'' to the compilation to get
  the compiler output to tell you where the problems are.

\end{itemize}

%-----------------------------------------------------------------------------
\subsection{PGI}

The general feeling with these compilers is that they produce slower
code than the PathScale compilers.

\ \\
Machine-specific notes:
\begin{itemize}

\item (on Kraken) The bounds checking routines do not like the way some of
the arrays are allocated and flag them as errors.  For instance, in
{\tt layout.f90}, we allocate a box as:
\begin{verbatim}
    type(box) :: bxs(3**get_dim(b))
\end{verbatim}
but the bounds checking routine seems to fail on this.

\item (on hopper) There are problems with code compiled with {\tt -fast}
instead of just {\tt -O}.  gcc/gfortran seem to work better and 
are only slighly slower.
\end{itemize}