### 2012-05-22 docs/makefiles/makefiles.tex
finish for now

\label{ch:make}

\section{Build Process Overview}

The \maestro\ build system uses features of GNU make, which is
typically the default on systems.  The \maestro\ executable is built
in the problem's directory (one of the directories under {\tt
SCIENCE/}, {TEST\_PROBLEMS}, or {UNIT\_TESTS}).  This directory will
contain a makefile, {\tt GNUmakefile}, that includes all the necessary
information to build the executable.

The main macros that define the build process are split across several
files.  The 4 main files are:
\begin{itemize}
\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakedefs.mak}:

  This setups the basic macros, includes the options for the selected
  compiler, builds the list of object and source files, and defines
  the compile and link command lines.

\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakeMPI.mak}:

  This implements any changes to the compiler names and library
  locations necessary to build a parallel executable with MPI.

\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakerules.mak}:

  This creates the various build targets and specifies the rules for
  building the object files, the list of dependencies, and some other
  lesser-used targets (tags for editors, documentation, etc.)

\item {\tt MAESTRO/GMaestro.mak}:

  This is a \maestro-specific file that gathers all of the various
  modules that are used to build a typical \maestro\ application
  and integrates with the \boxlib\ build system.  Every \maestro\
  problem's {\tt GNUmakefile} will include this file.

\end{itemize}

\maestro\ gets the location of the \boxlib\ library through the 
{\tt BOXLIB\_HOME} variable.  This should be set as an environment
variable in your shell start-up files (e.g.\ {\tt .bashrc} or {\tt
.cshrc}).

The \boxlib\ build system separates the compiler-specific information
from the machine-specific information---this allows for reuse of the
compiler information.  The only machine-specific parts of the build system
are for the MPI library locations, contained in {\tt GMakeMPI.mak}.
The compiler flags for the various compilers are listed in the
files in {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/comps/}.  The compiler
is set via the {\tt COMP} variable in the problem's {\tt GNUmakefile}.

There are several options in addition to the compiler that affect the
build process: {\tt MPI}, {\tt OMP}, and {\tt NDEBUG}---these turn on/off
MPI parallelization, OpenMP parallelization, and debugging.  Together,
these choices along with the compiler name are reflected in the name
of the executable.

When the `{\tt make}' command is run, the object and module files are
written into a directory {\tt t/{\em OS}.{\em COMP}.{\em other}/},
where {\tt \em OS} is the operating system detected by the build
system, {\tt \em COMP} is the compiler used, and {\tt \em other}
reflects any other build options (MPI, OpenMP, debugging, etc.) used.
Separating each build into a separate subdirectory under the problem
directory allows for multiple builds of \maestro\ to exist
side-by-side in the problem directory.

\subsection{Finding Source Files}

The \maestro\ executable is built from source distributed across a
number of directories.  In each of these directories containing source
files, there is a {\tt GPackage.mak} file.  This file has a number of
lines of the form:
\begin{verbatim}
f90sources += file.f90
\end{verbatim}
where {\tt file.f90} is a source file that should be built when this
directory is added to the list of build directories.  For old
fixed-form Fortran files, the files should be added to the {\tt
fsources} variable instead of {\tt f90sources}.

The \boxlib\ build system relies on the vpath functionality of {\tt
make}.  In a makefile, the {\tt vpath} variable holds search path used
to locate source files.  When a particular file is needed, the
directories listed in {\tt vpath} are searched until the file is
found.  The first instance of a file is used.  We exploit this feature
by always putting the build directory first in {\tt vpath} (this is
done in {\tt GMakerules.mak}).  This means that if a source file is
placed in the build directory, that copy will override any other
version in the source path.

In \maestro, the {\tt vpath} variable is set using the macros defined
in {\tt GMaestro.mak}.  A user does not need to set this variable
explicitly.  Additional source locations are added in the manner
described below (see \S~\ref{sec:make:otherfiles}).

\subsection{Dependencies}

There is no need to explicitly define the dependencies between the
source files for Fortran modules.  The scripts in {\tt
BOXLIB\_HOME/Tools/F\_scripts/} are run at the start of the build
process and parse all the source files and make an explicit list of
the dependency pairs.  This process is defined in {\tt
GMakerules.mak}.

A few files use explicit `{\tt include}' statements to include Fortran
source in other source files.  Any include file locations should be
added to {\tt Fmincludes} variable in the problem's {\tt GNUmakefile}.
This does not occur frequently.  For the case of the {\tt helmeos}
equation of state, this is done automatically in {\tt GMaestro.mak}.

\subsection{Files Created at Compile-time}

Several files are created at build-time:
\begin{itemize}
\item {\tt probin.f90}:

  This is the module that controls runtime parameters.  This is
  created by the script
  {\tt write\_probin.py} in {\tt \$\{BOXLIB\_HOME\}/Tools/F\_scripts/}.  The
  makefile logic for creating it is in {\tt GMaestro.mak}.  At compile
  time, the problem, main {\tt MAESTRO/}, and {\tt extern/}
  directories are searched for {\tt \_parameter} files.  These files
  are parsed and the {\tt probin.f90} file is output containing the 
  runtime parameters, their default values, and the logic for reading
  and interpreting the inputs file.

\item {\tt build\_info.f90}:

  This is a module that contains basic information about the build
  environment (compiler used, build machine, build directory, compiler
  flags, etc.).  This is created by the script {\tt make\_build\_info}
  in {\tt \$\{BOXLIB\_HOME\}/Tools/F\_scripts/} from {\tt
  GMaestro.mak} by passing in a number of makefile variables.  This is
  rewritten everytime the code is compiled.  The primary use of this
  module is writing out the {\tt job\_info} file in the plotfiles.

\item ({\tt network.f90}):

  This is generated at compile time {\em only} for the {\tt
  general\_null} network.  The {\tt general\_null} network allows the
  user to define a list of non-reacting species builds the {\tt
  network.f90} based on this list.  The makefile logic for building
  the {\tt network.f90} is in the {\tt GPackage.mak} in {\tt
  extern/networks/general\_null}.  The script {\tt write\_network.py}
  in that directory does the actual parsing of the species file and
  outputs the {\tt network.f90}.

\end{itemize}

\section{\maestro\ Problem Options}

\subsection{Problem-specific Files}
\label{sec:make:otherfiles}

If a problem has a unique file that is needed as part of the build,
then that file should be added to a {\tt GPackage.mak} file in the
problem's directory.  Since, by default, problems don't have a {\tt
GPackage.mak}, the build system needs to be told to look in the
problem directory for unique sources.  This is accomplished by adding
the problem's directory to the {\tt EXTRA\_DIR} variable in the
problem's {\tt GNUmakefile}.

Note that this is not necessary if you place a custom version of 
a source file in the problem's directory.  Since that file is already
listed in the {\tt GPackage.mak} in its original location, the build
system will know that it needs to be built.  Since the {\tt vpath}
variable puts the problem's directory at the start of the search
path, the version of the file in the problem's directory will be 
found first.

\subsection{Defining EOS, Network, and Conductivity Routines}

Each \maestro\ problem needs to define an equation of state, a
reaction network, and a routine to compute the conductivities (for
thermal diffusion).  This is true even if the problem is not doing
reactions of thermal diffusion.  These definitions are specified
in the problem's {\tt GNUmakefile}.

\begin{itemize}
\item {\tt EOS\_DIR}:

  This variable points to the path (relative to {\tt fParallel/}) of
  the equation of state used for the build.  Choices that work
  with \maestro\ are:
  \begin{itemize}
  \item {\tt extern/EOS/helmeos}
  \item {\tt extern/EOS/gamma\_law\_general}
  \end{itemize}

\item {\tt CONDUCTIVITY\_DIR}:

  This variable points to the conductivity routine used for the build
  (relative to {\tt fParallel/}).  Choices that work with \maestro\
  are:
  \begin{itemize}
  \item {\tt extern/conductivity/constant}
  \item {\tt extern/conductivity/timmes\_stellar}
  \end{itemize}
  If diffusion is not being used for the problem, this should be set
  to {\tt extern/conductivity/constant}.

\item {\tt NETWORK\_DIR}:

  This variable points to the reaction network used for the build
  (relative to {\tt fParallel/}).  A lot of options are present in
  {\tt extern/networks/}.  A network is required even if you are not
  doing reactions, since the network defines the species that are
  advected and interpreted by the equation of state.

  A special choice, {\tt extern/networks/general\_null} is a general
  network that simply defines the properties of one or more species.
  This requires an inputs file, specified by {\tt
  GENERAL\_NET\_INPUTS}.  This inputs file is read at compile-time and
  used to build the {\tt network.f90} file that is compiled into the
  source.

\end{itemize}

\subsection{Core \maestro\ modules}

Several modules are included in all \maestro\ builds by default.
From \boxlib, we alway include:
\begin{itemize}
\item {\tt \$\{BOXLIB\_HOME\}/Src/F\_BaseLib}
\item {\tt \$\{BOXLIB\_HOME\}/Src/LinearSolvers/F\_MG}
\end{itemize}

\noindent From {\tt extern}, we always include
\begin{itemize}
\item {\tt extern/model\_parser}
\item {\tt extern/LAPACK}
\item {\tt extern/random}
\item {\tt extern/BLAS}
\item {\tt \$\{EOS\_DIR\}}
\item {\tt \$\{NETWORK\_DIR\}}
\item {\tt \$\{CONDUCTIVITY\_DIR\}}
\end{itemize}
for most networks, we automatically add in {\tt extern/VODE} as well.

\noindent From {\tt MAESTRO/}, we add
\begin{itemize}
\item {\tt MAESTRO/constants}
\item {\tt MAESTRO/Source}
\end{itemize}
(although see the unit tests section below regarding {\tt MAESTRO/Source}.

\noindent Finally, any extra directories listed in the {\tt EXTRA\_DIR}
variable are included.

For each of these included directories, {\tt GMaestro.mak} adds the list of source files
defined in their {\tt GPackage.mak} to the list of files to be compiled.
It also adds each of these directories to the {\tt vpath} as a directory
for the build process to search in for source files.

\subsection{Unit Tests}

Sometimes we only want to use a few of the standard \maestro\
routines, for example in a unit test where we are testing only a small
part of the \maestro\ algorithm indepenedently.  In this case, we
don't want to comple all of the files in {\tt MAESTRO/Source}.  If we
set {\tt UNIT\_TEST := t} in our problem's {\tt GNUmakefile}, then the
{\tt GPackage.mak} in {\tt MAESTRO/Source} is not read, so those files
are not automatically put into the list of files to compile.  Instead,
the problem should create its own {\tt GPackage.mak} listing only the
subset of files that are to be compiled.  {\tt MAESTRO/Source} is put
into the {\tt vpath} search path for sources, so those files will
still be found as needed.

\section{Special Targets}

\subsection{Debugging ({\tt print-*})}

To see the contents of any variable in the build system, you can build
the special target {\tt print-{\em varname}}, where {\tt {\em
varname}} is the name of the variable.  For example, to see what the
Fortran compiler flags are, you would do:
\begin{verbatim}
make print-FFLAGS
\end{verbatim}
This would give (for {\tt gfortran}, for example):
\begin{verbatim}
FFLAGS is -Jt/Linux.gfortran/m -I t/Linux.gfortran/m -O2 -fno-range-check
\end{verbatim}
This functionality is useful for debugging the makefiles.

\subsection{{\tt clean} and {\tt realclean}}

Typing `{\tt make clean}' deleted the object and module files for the
current build (i.e., the current choice of {\tt MPI}, {\tt NDEBUG},
{\tt COMP}, and {\tt OMP}).  This also removes any of the compile-time
generated source files.  Any other builds are left unchanged.

Typing `{\tt make realclean}' deletes the object and module files for
all builds---i.e., the entire {\tt t/} directory is removed.

\section{Extending the Build System}

\subsection{Adding a Compiler}

Properties for different compilers are already defined in 
{\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/comps/}.
Each compiler is given its own file.  The appropriate file is included
into {\tt GMakedefs.mak} by looking at the {\tt COMP} variable and the
operating system.  These compiler files define the compiler flags for
both optimized and debug compiling.  Additionally, the variable {\tt
FCOMP\_VERSION} should be defined there, based on the output from the
compiler, to provide the compiler version for output into the {\tt
job\_info} file at runtime.

\subsection{Parallel (MPI) Builds}

When building with MPI, the build system needs to know about the location of the
MPI libraries.  If your local MPI has the {\tt mpif90} and {\tt mpicc} wrappers
installed and working, then \maestro\ can use these simply by setting the
environment variable:
\begin{verbatim}
BOXLIB_USE_MPI_WRAPPERS=1
\end{verbatim}
Otherwise, you will need to edit {\tt GMakeMPI.mak} and add a section
specific to your machine with the compiler and library location.  It
is best to simply copy an existing similar portion of the makefile and
adjust it to your system.  Most national supercomputing facilities are
already supported, and parallel builds on them should work out of the
box.