### 2012-06-05 docs/preface/preface.tex

This documentation is a work in progress.  \maestro\ is rapidly
evolving, so portions of the text are likely to be incomplete or
out-of-date. \\

\noindent \maestro\ is developed at the Center for Computational Sciences and
Engineering at Lawrence Berkeley National Laboratory and at Stony
Brook University.  Development was supported by the SciDAC Program of
the DOE Office of Mathematics, Information, and Computational Sciences
and the SciDAC Program of the DOE Office of High Energy Physics under
the U.S. Department of Energy under contract No.\ DE-AC02-05CH11231 to
LBNL and by a DOE/Office of Nuclear Physics grant No.\
DE-FG02-06ER41448, to Stony Brook.  \\

\noindent \maestro\ uses the \boxlib\ library,
developed at the Center for Computational Sciences and
Engineering at Lawrence Berkeley National Laboratory.  \boxlib\
is distributed separately from \maestro\ (see https:/\hspace{-0.25em}/ccse.lbl.gov/BoxLib/ )
\\

\noindent The {\tt helmeos} equation of state was developed by Frank Timmes (ASU).
The version in \maestro\ was downloaded from his website and the
interface changes slightly for threadsafeness.  We thank Frank for
making it available.

\noindent The current version of the \maestro\ User's Guide can be found in 
the \maestro\ source tree under {\tt fParallel/MAESTRO/docs/}.